Notes on area-length problem for CLE

### 2014-03-15 area-length/notes.tex
Some text on pinning operator and Hausdorff dimension

\subsection{Self-avoiding polygons}
\label{sec:self-avoid-polyg}

@@

@@

Local observables w.r.t. SLE can be described by correlation functions of boundary conformal field
theory. So we need a theory which reproduces CFT in the limit $\mu,\mu_{B}\to 0$ and has non-local
observables such as area and length.

@@

@@

@@

In theory with the action
$$S=\frac{1}{4\pi} \int \sqrt{g}\left( \frac{1}{2} g^{\mu\nu} \partial_{\mu} \varphi \partial_{\nu}
  \varphi + Q R\varphi + 4\pi\mu e^{2b\varphi}\right) d^{2}x,$$ 
where $g$ is a background metric we need to assume some regularization procedure to define a
functional integral, because we want to integrate over measure that does not depend upon metric.

The idea is that the form of the action is the same but parameters are changed. 
So now $\mu$ is not a classical cosmological constant and $\mu,b,Q$ are connected through
consistency requirements for the theory. 
The consistency requirement is background-independence. The action should not change when we choose
another background metric and shift field $\varphi$ accordingly. 
Let $\tilde g = e^{\sigma} g$, redefine $\varphi$ as $\tilde \varphi=\varphi + \frac{Q}{2}\sigma$,
then all the terms with $\sigma$ are cancelled if we take into account contributions from matter
partition function. But it is possible only if 
$$Q=\frac{25-c_{M}}{6}$$. 
To make exponential term background-independent we need to have $Q=\frac{1}{b}+b$.

@@

>>>>

v\subsection{Self-avoiding polygons}
\label{sec:self-avoid-polyg}

@@

Using their numeric results, authors of the paper \cite{richard2001scaling} were able to come with
the following scaling for rooted generating function:
\begin{equation}
  \label{eq:112}
  G^{(r)}(\mu,\mu_B)\sim \mu^{\frac{1}{3}} \frac{\mathrm{Ai}'(s)}{\mathrm{Ai}(s)},\quad \mbox{where}\; s=\left(\frac{\mu_{B}^{2}}{\mu}\right)^{\phi}
\end{equation}
We want to get physical and geometrical understanding of this result and its generalization to
different measures on closed curves.

@@

Local observables w.r.t. SLE can be described by correlation functions of boundary conformal field
theory. So we need a theory which reproduces CFT in the limit $\mu,\mu_{B}\to 0$ and has non-local
observables such as area and length. Note also that the parameter $\kappa$ is connected with the
central charge of conformal field theory by the relation
\begin{equation}
  \label{eq:113}
  c=\frac{(8-3\kappa)(\kappa-6)}{2\kappa}
\end{equation}

@@

Hausdorff dimension in CFT is calculated by the insertion of certain primary field called ``pinning
operator'' \cite{bettelheim2005harmonic}. Expanded presentation can be found in the paper
\cite{gruzberg2006stochastic}. The argument is as follows.

Consider a closed curve. Harmonic measure of some part of the curve can be understood as a charge of
that part if the curve is made of conducting material and its total charge is equal to one. 
Closed curve can be covered by small circles of radius $r$ centered at points
$z_{i}$. Consider one such circle. The curve divides the circle into two parts, inner and outer
w.r.t. to curve. Outer part can be conformally mapped to the unit disk. Denote this map by $w(z)$.
With the normalization such that one point is fixed and derivative at infinity is equal to one the
moments of harmonic measure in small circle of radius $r$ scale as
\begin{equation}
  \label{eq:116}
  \mu(r)\sim |w(r)|\sim r |w'(r)|
\end{equation}
So scaling of the moments of $\mu(r)$ are determined by scaling of $|w'(r)|$:
\begin{equation}
  \label{eq:117}
  \prec |w'(r)|^{h}\succ \sim r^{\Delta(h)}, \quad r\to 0
\end{equation}

The moments of harmonic measure for the whole curve can be written as a sum over small circles:
\begin{equation}
  \label{eq:120}
  M_{h}=\sum_{i=1}^{N} \mu(z_{i},r)^{h}
\end{equation}

In the limit $r\to 0$ number of circles grows and moments scale as
\begin{equation}
  \label{eq:121}
  M_{h}\sim r^{\tau(h)}
\end{equation}
The function $\tau(h)$ is called multifractal spectrum. For simple curve $\tau(h)=h-1$, so anomalous
exponents $\delta(h)$ are defined as $\delta(h)=\tau(h)-h+1$.

Note that $M_{0}$ is a total number of circles and $\tau(0)=-d_{H}$ by definition of Hausdorff
dimension. For $h=1$ the sum is equal to the total charge of the curve so it does not scale and
$\tau(1)=0$.

To calculate multifractal spectrum we need to average over the values at points $z_{i}$. One can
assume the ergodicity and rewrite $M_{h}$ as
\begin{equation}
  \label{eq:122}
  M_{h}=N \prec \mu(z,r)^{h}\succ
\end{equation}
for some point $z$. The total number of circles $N$ is equal to zeroth momentum $M_{0}$, so
$M_{h}\sim r^{\tau(0)} \prec \mu(z,r)^{h}\succ$. Local multifractal spectrum $\tilde\tau(h)$ is
defined as $ \prec \mu(z,r)^{h}\succ\sim r^{\tilde\tau(h)}$ and local anomalous exponents are
$\Delta_{bulk}(h)=\tilde\tau(h) -h$.

Following relations hold for multifractal spectrum and anomalous exponents:
$\tilde\tau(h)=\tau(h)-\tau(0)$, $\Delta_{bulk}(h)=\delta(h)-\delta(0)$, $d_{H}=1+\Delta_{bulk}(1)$

Multifractal spectrum and anomalous exponents were first obtained by Duplantier in paper
\cite{duplantier2000conformally} using quantum gravity. The result is
\begin{equation}
  \label{eq:123}
  \Delta_{bulk}(h)=-\frac{h}{2}+\left(\frac{1}{16}+\frac{1}{4\kappa}\right)
  (\kappa-4+\sqrt{(\kappa-4)^{2}+16\kappa h})
\end{equation}
\begin{equation}
  \label{eq:124}
  \tau(h)=\frac{1}{2}(h-1) +\frac{25-c}{24}\left(\sqrt{\frac{24 h +1-c}{25-c}}-1\right), 
\end{equation}
where $c$ and $\kappa$ are connected by \eqref{eq:113}.

So we have following result for Hausdorff dimension (eq \eqref{eq:5}):
\begin{equation}
  \label{eq:125}
  d_{H}=1+\frac{\kappa}{8}
\end{equation}

The stochastic averages for (local) moments of harmonic measure are represented by different CFT
correlation functions depending upon the position of point of interest. If we are interested in a
bulk point lying on a curve, we need to put at this point a special operator. It is $\Psi_{10}$
which has an interpretation as an operator that creates two lines coming from the point in the bulk.
To see it one can consider circle of small radius $\varepsilon$ centered at $z$. Put some conformal
boundary condition on one half of the circle and another condition on the rest of the circle. Then
critical curve would join the points of change in boundary conditions. The expectation values of
observables in this geometry are given by CFT correlation functions with two boundary condition
changing operators inserted:
\begin{equation}
  \label{eq:119}
  \left< \Psi_{12}(z')\Psi_{12}(z'')\right>
\end{equation}
where $|z'-z|\leq \varepsilon$ and $|z''-z|\leq \varepsilon$. The boundary condition changing
operators in Coulomb gas picture can be presented as the vertex operators
$V_{\alpha_{12}}=e^{i\sqrt{2}\alpha_{12}\varphi(z)}$. When $\varepsilon\to 0$ we can fuse this two
operators to single operator $V_{2\alpha_{12}}$. Note that $\alpha_{12}=-\frac{1}{2}\alpha_{-}$
\eqref{eq:55}, so $V_{2\alpha_{12}}\sim V_{-\alpha_{-}}\sim V_{\alpha_{+}}$ (?)

The moments of harmonic measure are connected with the CFT correlation functions by
\begin{equation}
  \label{eq:118}
  \left<O_{h}(r) \Psi_{10}(0)\Psi(\infty)\right>\sim r^{h'-2h}\prec |w'(r)|^{h}\succ
\end{equation}

@@

Let $D$ denote $\mathbb{C}$, $\mathbb{C}^{*}$ or $\mathbb{D}$.
\begin{theorem}
Every Riemann surface $M$ is conformally equivalent to the quotient $D/\Gamma$ with $\Gamma$  a freely acting
discontinuous group of fractional transformations preserving $D$.
\end{theorem}

Let us consider the case of Riemann surfaces with universal covering $\mathbb{H}\sim \mathbb{D}$ and
denote by $J_{H}$ the complex analytic covering $J_{H} : \mathbb{H}\to M$ . In this case $\Gamma$
(the automorphism group of $J_{H}$ ) is a finitely generated Fuchsian 
group $\Gamma\subset PSL(2,
\mathbb{R}) = SL(2, \mathbb{R})/{I, -I}$ acting on $\mathbb{H}$ by linear fractional transformations
\begin{equation}
  \label{eq:106}
  w\in \mathbb{H}\quad \gamma\cdot w=\frac{aw+b}{cw+d}\in\mathbb{H},\quad
  \gamma=\begin{pmatrix} a & b\\ c & d\end{pmatrix} \in\Gamma\subset PSL(2,\mathbb{R})
\end{equation}

By the fixed point equation
\begin{equation}
  \label{eq:107}
  w_{\pm}=\frac{a-d\pm\sqrt{(a+d)^{2}-4}}{2c}
\end{equation}
it follows that $\gamma\neq I$ can be classified according to the value of $|\mathrm{tr} \gamma|$:
\begin{enumerate}
\item Elliptic case: $|\mathrm{tr} \gamma|<2$, $\gamma$ has one fixed point on $\mathbb{H}$:
  $w_{-}=\bar w_{+}\not\in\mathbb{R}$ and $M$ has branch point
\item Parabolic case: $|\mathrm{tr}\gamma|=2$, $w_{-}=w_{+}\in\mathbb{R}$ and Riemann surface $M$
  has a puncture.
\item Hyperbolic case: $|\mathrm{tr}\gamma|>2$, Riemann surface has handles. 
\end{enumerate}

   A Riemann surface $M$ isomorphic to the quotient $\mathbb{H}/\Gamma$ has the Poincar\'e metric $\hat g$ as the
unique metric with scalar curvature $R_{\hat g} = -1$ compatible with its complex structure. This
implies the uniqueness of the solution of the Liouville equation on $M$. The Poincar\'e metric
on $\mathbb{H}$ is
\begin{equation}
  \label{eq:108}
  ds^{2}=\frac{|dw|^{2}}{(\Im w)^{2}}
\end{equation}
Note that $P SL(2, \mathbb{R})$ transformations are isometries of $\mathbb{H}$ endowed with the Poincar\'e metric.
   An important property of $\Gamma$ is that it is isomorphic to the fundamental group $\pi_{1}(M)$.

Substitute $w=J_{H}^{-1}(z)$ in \eqref{eq:108} and get Poincar\'e metric on $M$:
\begin{equation}
  \label{eq:109}
  ds^{2}=2\hat g_{z\bar z}|dz|^{2}=e^{\varphi(z,\bar z)} |dz|^{2}=\frac{|J_{H}^{-1}(z)'|^{2}}{(\Im
    J_{H}^{-1}(z))^{2}} |dz|^{2}
\end{equation}

The curvature in complex coordinates is given by
\begin{equation}
  \label{eq:110}
  R_{\hat g}=-\hat g^{z\bar z}\partial \bar \partial \ln \hat g_{z\bar z}, \quad \hat g^{z\bar z}=2e^{-\varphi}
\end{equation}
The condition $R_{\hat g}=-1$ is equivalent to Liouville equation
\begin{equation}
  \label{eq:111}
  \partial \bar \partial \varphi(z,\bar z)=\frac{1}{2} e^{\varphi(z,\bar z)}
\end{equation}
The field $\tilde\varphi=\varphi+\ln \mu, \; \mu>0$ defines a metric of constant curvature $-\mu$.

@@

In theory with the action
$$S=\frac{1}{4\pi} \int \sqrt{g}\left( \frac{1}{2} g^{\mu\nu} \partial_{\mu} \varphi \partial_{\nu}
  \varphi + Q R\varphi + 4\pi\mu e^{2b\varphi}\right) d^{2}x,$$ 
where $g$ is a background metric, we need to assume some regularization procedure to define a
functional integral, because we want to integrate over measure that does not depend upon metric
\begin{equation}
  \label{eq:114}
  ||\delta\varphi||^{2}=\int \sqrt{g} e^{\varphi(x)}(\delta\varphi)^{2} d^{2}x
\end{equation}
This measure is non-linear and non translation-invariant and we need to substitute it with the
measure
\begin{equation}
  \label{eq:115}
    ||\delta\varphi||^{2}=\int \sqrt{g}(\delta\varphi)^{2} d^{2}x,
\end{equation}
that is  w.r.t. to change $\varphi(x)\to \varphi(x) +C(x)$ for arbitrary fixed function $C(x)$.

There is for now no way to do it, so 
the idea is that the form of the action after regularization is the same but parameters are changed.
The parameters are then obtained from consistency condition, which is background-independence of the
theory.

So now $\mu$ is not a classical cosmological constant and $\mu,b,Q$ are connected through
consistency requirements for the theory. 
The consistency requirement is background-independence. The action should not change when we choose
another background metric and shift field $\varphi$ accordingly. 
Let $\tilde g = e^{\sigma} g$ and redefine $\varphi$ as $\tilde \varphi=\varphi +
\frac{Q}{2}\sigma$.

Consider first the case $\mu=0$.
then all the terms with $\sigma$ are cancelled if we take into account contributions from matter
partition function. But it is possible only if 
$$Q=\frac{25-c_{M}}{6}$$. 
To make exponential term background-independent we need to have $Q=\frac{1}{b}+b$.

@@

\section{Scaling of generating function for polygons}
\label{sec:scal-gener-funct}

The solution is
\begin{equation}
  \label{eq:105}
  \left< V_{\alpha_{01}} B_{\beta_{10}}\right>
\end{equation}
